---
title: "diamond"
author: "John King"
date: "4/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project

```{r, echo=FALSE, warning=FALSE, message='false', results='hide'}
library(dplyr)
diamond = read.csv('./diamonds.csv')
diamond$clarity = as.factor(diamond$clarity)
new.levels = c('below', 'below' , 'below', 'above', 'below', 'below', 'below', 'above', 'below', 'above', 'above')
x <- factor(new.levels[diamond$clarity])
diamond$claritybool = x
#diamond = diamond %>% filter(diamond$total_sales_price < 10000)

#Convert everything to factor
diamond$shape = as.factor(diamond$shape)
diamond$color = as.factor(diamond$color)
diamond$fancy_color_dominant_color = as.factor(diamond$fancy_color_dominant_color)
diamond$fancy_color_secondary_color = as.factor(diamond$fancy_color_secondary_color)
diamond$fancy_color_overtone = as.factor(diamond$fancy_color_overtone)
diamond$fancy_color_intensity = as.factor(diamond$fancy_color_intensity)


diamond$cut = as.factor(diamond$cut)
diamond$symmetry = as.factor(diamond$symmetry)
diamond$polish = as.factor(diamond$polish)

diamond$girdle_min = as.factor(diamond$girdle_min)
diamond$girdle_max = as.factor(diamond$girdle_max)
diamond$culet_size = as.factor(diamond$culet_size)
diamond$culet_condition = as.factor(diamond$culet_condition)
diamond$fluor_color = as.factor(diamond$fluor_color)
diamond$fluor_intensity = as.factor(diamond$fluor_intensity)
diamond$lab = as.factor(diamond$lab)
diamond$eye_clean = as.factor(diamond$eye_clean)
diamond$widthLength = diamond$meas_width * diamond$meas_length

#str(diamond)

diamond_clean = diamond %>% dplyr::select(-c(diamond_id, clarity, date))

#for convenience
#diamond_clean = diamond_clean %>%
#  dplyr::select(-c(fancy_color_dominant_color, fancy_color_secondary_color, #fancy_color_overtone, fancy_color_intensity))

#diamond_clean = diamond_clean %>% 
#  dplyr::select(-color)

diamond_clean = diamond %>% dplyr::select(c(widthLength, depth_percent, table_percent, meas_depth, total_sales_price, claritybool))

#diamond_clean = diamond %>% dplyr::select(c(size, depth_percent, table_percent, meas_width, meas_length, meas_depth, total_sales_price, claritybool))

```


```{r, echo=FALSE, warning=FALSE, message='false', results='hide'}
#barplot(dplyr::select(c()))

```



```{r, echo=FALSE, warning=FALSE, message='false', results='hide'}
#library(MASS)
#train_index = sample(1:nrow(diamond_clean), nrow(diamond_clean) / 2, rep=FALSE)

#train = diamond_clean[train_index, ]
#test = diamond_clean[-train_index, ]

#lda.fit = lda(claritybool~., data = train)
#lda.fit = lda(train[, -20], grouping = train[,20])
#lda.pred = predict(lda.fit, test[, -20])
#names(lda.pred)

```


```{r, echo=FALSE, warning=FALSE, message='false', results='hide'}
library(class)
library(caret)

#ls()

set.seed(780)

index.train = sample(1:dim(diamond_clean)[1], dim(diamond_clean)[1], replace=FALSE)
index.test = -index.train
#index.test = sample(1:dim(diamond_clean)[1], 000, replace=FALSE)
```


```{r, echo=FALSE, warning=FALSE, message='false', results='hide'}
training.y = diamond_clean$claritybool[index.train]
training.x = diamond_clean[index.train, ] %>% dplyr::select(-claritybool)

flds = createFolds(training.y, k = 10, list = TRUE, returnTrain = FALSE)

K = c(10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)

cv_error = matrix(NA, 10, 11)

for(j in 1:11) {
  k = K[j]
  for(i in 1:10) {
    test_index = flds[[1]]
    testX = training.x[test_index, ]
    trainX = training.x[-test_index, ]
    
    trainY = training.y[-test_index]
    testY = training.y[test_index]
    
    knn.pred = knn(trainX, testX, trainY, k = k)
    cv_error[i, j] = mean(testY != knn.pred)
  }
}

apply(cv_error, 2, mean)
```

```{r, echo=FALSE, warning=FALSE, message='false', results='hide'}
set.seed(776)
index.sub = sample(1:dim(diamond_clean)[1], dim(diamond_clean)[1], replace=FALSE)

diamond_sub = diamond_clean[index.sub, ]

index.train = sample(1:dim(diamond_sub)[1], dim(diamond_sub)[1] / 2, replace=FALSE)
index.test = -index.train
#index.test = sample(1:dim(diamond_clean)[1], 000, replace=FALSE)
```

```{r, echo=FALSE, warning=FALSE, message='false'}
training.y = diamond_sub$claritybool[index.train]
training.x = diamond_sub[index.train, ] %>% dplyr::select(-claritybool)
testing.x = diamond_sub[index.test, ] %>% dplyr::select(-claritybool)
testing.y = diamond_sub$claritybool[index.test]

knn.pred = knn(training.x, testing.x, training.y, k = 11)
table(knn.pred, testing.y)
```
New:

Remove High Cost Diamonds k = 17

testing.y
knn.pred above below
   above 25441 16336
   below 19668 32954
   
Misclassification rate: 38.14%

Remove High Cost of Diamonds k = 11
testing.y
knn.pred above below
   above 25661 16431
   below 19448 32859
   
Misclassification rate: 38.01%


Full dataset k = 17

testing.y
knn.pred above below
   above 27661 18735
   below 24261 39195
   
Misclassification rate: 39.14%

k = 11

testing.y
knn.pred above below
   above 28190 19247
   below 23732 38683
   
Misclassification rate: 39.12%

Cleaning k = 11

testing.y
knn.pred above below
   above 30892 14487
   below 21030 43443
   
Misclassification rate: 32.33%



OLD:

Full
        testing.y
knn.pred above below
   above 27761 18939
   below 24161 38991

Remove High cost >= 10000
 testing.y
knn.pred above below
   above 25059 16078
   below 20149 33113

```{r, echo=FALSE, warning=FALSE, message='false', results='hide'}

set.seed(1444)
index.sub = sample(1:dim(diamond_clean)[1], dim(diamond_clean)[1], replace=FALSE)

diamond_sub = diamond_clean[index.sub, ]

index.train = sample(1:dim(diamond_sub)[1], dim(diamond_sub)[1] / 2, replace=FALSE)
index.test = -index.train
#index.test = sample(1:dim(diamond_clean)[1], 000, replace=FALSE)
```

```{r, echo=FALSE, warning=FALSE, message='false'}
training.y = diamond_sub$claritybool[index.train]
training.x = diamond_sub[index.train, ] %>% dplyr::select(-claritybool)
testing.x = diamond_sub[index.test, ] %>% dplyr::select(-claritybool)
testing.y = diamond_sub$claritybool[index.test]

knn.pred = knn(training.x, testing.x, training.y, k = 11)
table(knn.pred, testing.y)
```
NEW:

Remove High Cost Diamonds k = 17

testing.y
knn.pred above below
   above 25260 16243
   below 19780 33116
   
Misclassification rate: 38.16%

k = 11

testing.y
knn.pred above below
   above 25767 16652
   below 19273 32707
   
Misclassification rate: 38.06% 

Full Dataset k = 17

testing.y
knn.pred above below
   above 27617 18995
   below 24222 39018
   
Misclassification rate: 39.34%

k = 11

testing.y
knn.pred above below
   above 28365 19358
   below 23474 38655
   
Misclassification rate: 39.00%

Cleaning k = 11

testing.y
knn.pred above below
   above 31067 14698
   below 20772 43315
   
Misclassification rate: 32.29% 



OLD:

Full
testing.y
knn.pred above below
   above 27761 19225
   below 24078 38788

Rmove High Cost (>=10000)
testing.y
knn.pred above below
   above 25417 16300
   below 19752 32930
   
0.6180892
